A Parallel, Backjumping Subgraph Isomorphism Algorithm using Supplemental Graphs

### 2015-06-22 paper/paper.tex
Wording tweaks

A \emph{non-induced subgraph isomorphism} is an injective mapping $i : P \rightarrowtail T$ from a
graph $P$ to a graph $T$ which preserves adjacency---that is, if $v \sim_{P} w$ then we require
$i(v) \sim_{T} i(w)$ (and thus if $v$ has a loop, then $i(v)$ must have a loop). The
\emph{non-induced subgraph isomorphism problem} is to find such a mapping from a given pattern graph
$P$ to a given target graph $T$.  (The \emph{induced subgraph isomorphism problem} additionally
requires that if $v \not\sim_{P} w$ then $i(v) \not\sim_{T} i(w)$; we discuss only the non-induced
version in this paper. Both variants are \NP-complete.)

@@

A \emph{path} in a graph is a sequence of distinct vertices, such that each successive pair of
vertices are adjacent; we also allow a path from a vertex to itself, in which case the first and
last vertices in the sequence are the same (and there is a \emph{cycle}). The \emph{distance}
between two vertices is the length of a shortest path between them. We write $G^d$ for the graph
with vertex set $\operatorname{V}(G)$, and edges between $v$ and $w$ if the distance between $v$ and
$w$ in $G$ is at most $d$.  We introduce the notation $G^{\left[c, l\right]}$ for the graph with
vertex set $\operatorname{V}(G)$, and edges between vertices $v$ and $w$ (not necessarily distinct)
precisely if there are at least $c$ paths of length exactly $l$ between $v$ and $w$ in $G$. The
following proposition may easily be verified (by observing that injectivity means paths are
preserved):

@@

    \item The $\KwSty{search}$ function, parallelised without changes, could attempt to exit early
        due to backjumping. We rule out this possibility by refusing to pass knowledge to the left:
        that is, we do not allow speculatively-found backjumping conditions to change the return
        value of $\KwSty{search}$). This is for safety \cite{Trienekens:1990} and reproducibility:
        value-ordering heuristics can alter the performance of unsatisfiable instances when
        backjumping, and allowing parallelism to select a different backjump set could lead to an
        absolute slowdown \cite{Prosser:1993b}. To avoid this possibility, when a backjump condition
        is found, we \emph{must} cancel any speculative work being done to the right of its
        position, and \emph{cannot} cancel any ongoing work to the left.  This means that unlike in
        conventional backtracking search without learning, we should \emph{not} expect a linear
        speedup for unsatisfiable instances.

@@

Going against conventional wisdom, we saw that replacing strong inference with cheaper surrogates
could pay off, and that backjumping could be implemented cheaply enough to be beneficial. We also
saw parallelism give a substantial benefit. This was true even for relatively low runtimes, due to
us exploiting parallelism for pre-processing as well as for search.  Parallel backjumping has only
been given limited attention \cite{Conrad:1994,Habbas:1997,Cope:2000}.  However, a simple approach
has worked reasonably well here (in contrast to stronger learning systems, where successes in
parallelism appear to be rare).

>>>>

A \emph{non-induced subgraph isomorphism} is an injective mapping $i : P \rightarrowtail T$ from a
graph $P$ to a graph $T$ which preserves adjacency---that is, if $v \sim_{P} w$ then we require
$i(v) \sim_{T} i(w)$ (and thus if $v$ has a loop, then $i(v)$ must have a loop). The
\emph{non-induced subgraph isomorphism problem} is to find such a mapping from a given pattern graph
$P$ to a given target graph $T$.  (The \emph{induced subgraph isomorphism problem} additionally
requires that if $v \not\sim_{P} w$ then $i(v) \not\sim_{T} i(w)$, and variants also exist for
directed and labelled graphs; we discuss only the non-induced version in this paper. All these
variants are \NP-complete.)

@@

A \emph{path} in a graph is a sequence of distinct vertices, such that each successive pair of
vertices are adjacent; we also allow a path from a vertex to itself, in which case the first and
last vertices in the sequence are the same (and there is a \emph{cycle}). The \emph{distance}
between two vertices is the length of a shortest path between them. We write $G^d$ for the graph
with vertex set $\operatorname{V}(G)$, and edges between $v$ and $w$ if the distance between $v$ and
$w$ in $G$ is at most $d$.  We introduce the notation $G^{\left[c, l\right]}$ for the graph with
vertex set $\operatorname{V}(G)$, and edges between vertices $v$ and $w$ (not necessarily distinct)
precisely if there are at least $c$ paths of length exactly $l$ between $v$ and $w$ in $G$. The
following proposition may easily be verified by observing that injectivity means paths are
preserved:

@@

    \item The $\KwSty{search}$ function, parallelised without changes, could attempt to exit early
        due to backjumping. We rule out this possibility by refusing to pass knowledge to the left:
        that is, we do not allow speculatively-found backjumping conditions to change the return
        value of $\KwSty{search}$. This is for safety \cite{Trienekens:1990} and reproducibility:
        value-ordering heuristics can alter the performance of unsatisfiable instances when
        backjumping, and allowing parallelism to select a different backjump set could lead to an
        absolute slowdown \cite{Prosser:1993b}. To avoid this possibility, when a backjump condition
        is found, we \emph{must} cancel any speculative work being done to the right of its
        position, and \emph{cannot} cancel any ongoing work to the left.  This means that unlike in
        conventional backtracking search without learning, we should \emph{not} expect a linear
        speedup for unsatisfiable instances.

@@

Going against conventional wisdom, we saw that replacing strong inference with cheaper surrogates
could pay off, and that backjumping could be implemented cheaply enough to be beneficial. We also
saw parallelism give a substantial benefit. This was true even for relatively low runtimes, due to
us exploiting parallelism for pre-processing as well as for search.  Parallel backjumping has only
been given limited attention \cite{Conrad:1994,Habbas:1997,Cope:2000}.  However, a simple approach
has worked reasonably well here (in contrast to stronger clause-learning systems, where successes in
parallelism appear to be rare).
